### 2019-02-25 manuscript/methods.tex
some changes from "model-A/B" to "model+A/B"

% D/P ploidy model
% D/P-A/B ploidy hidden state model
% I/C breeding system model
% I/C-A/B breeding system and hidden state model
% ID/CD/CP ploidy and breeding system model
% ID/CD/CP-A/B ploidy, breeding system, and hidden state model

@@

Our second model assesses the signal of diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
Diversification rate differences explained by something (trait) other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}). %B: inserted "(trait)", because that seems to be the goal - a biological explanation?
In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
We call this the \textit{D/P--A/B ploidy and hidden state} model. 
We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
In this model polyploidization rate $\rho$ and diploidization rate $\delta$ are also included, and changes between hidden states are symmetrical with rate $\alpha$. %B: same dip vs. re-dip question, as above.

@@

To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
We call this BiSSE model the \textit{I/C breeding system} model. 
To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
We call this HiSSE model the \textit{I/C--A/B breeding system and hidden state model}.

@@

%B: What happens if SC->SI is not fixed? Asking for a friend.

@@

The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P--A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C--A/B was effectively caused by ploidy. %B: you may wish to get rid of "(or its correlates)", which I inserted; along with "caused". I know that causation is not strict here, but it clarifies the intended biological explanation. %B: was: "if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy."
There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
We refer to this as the \textit{ID/CD/CP--A/B} model.
A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 16 parameters.
The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states. Perhaps this complication is covered in the Discussion section.

% E todo: explain "pathway" method of robertson2011

@@

@@

Parameters for each of the 10 diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment. %B: fixme: I think the intent here is  "Parameter estimation [...] was peformed"
Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
We included a correction for incomplete sampling in all analyses, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models. %B: the number of models changes here?
For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots). % fixme: eventually, add supp info fig ref

@@

Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen? %B: fixme: my count too.
Six compared the models of polyploidy against each other (D/P and D/P--A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C--A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP--A/B, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
%R: Emma this is where the figure can be important to show that we have to make different input
% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily. I find it useful to imagine Spencer reading this, for example.

>>>>

% D/P ploidy model
% D/P+A/B ploidy hidden state model
% I/C breeding system model
% I/C+A/B breeding system and hidden state model
% ID/CD/CP ploidy and breeding system model
% ID/CD/CP+A/B ploidy, breeding system, and hidden state model

@@

Our second model assesses the signal of diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
Diversification rate differences explained by something (trait) other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}). %B: inserted "(trait)", because that seems to be the goal - a biological explanation?
In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
We call this the \textit{D/P+A/B ploidy and hidden state} model. 
We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
In this model polyploidization rate $\rho$ and diploidization rate $\delta$ are also included, and changes between hidden states are symmetrical with rate $\alpha$. %B: same dip vs. re-dip question, as above.

@@

To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
We call this BiSSE model the \textit{I/C breeding system} model. 
To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.

@@

%B: What happens if SC->SI is not fixed? Asking for a friend.  E: Surprisingly, Rosana reports, results are not turned on their head.

@@

The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P+A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C+A/B was effectively caused by ploidy. % B: I know that causation is not strict here, but it clarifies the intended biological explanation.
There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
We refer to this as the \textit{ID/CD/CP+A/B} model.
A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 16 parameters.
The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states. Perhaps this complication is covered in the Discussion section.

@@

% \subsection{Pathways to polyploidy} % E todo
% robertson2011 method, and then with diversification

@@

Parameters for each of the 10 diversification models were estimated using custom code in the RevBayes \citep{hoehna_2016} environment.
Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
We included a correction for incomplete sampling in all analyses, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models. %B: the number of models changes here?
For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots). % fixme: eventually, add supp info fig ref

@@

Using the marginal likelihood values, we calculated thirteen different Bayes factors.
Six compared the models of polyploidy against each other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
%R: Emma this is where the figure can be important to show that we have to make different input
% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily. I find it useful to imagine Spencer reading this, for example.
% E: "Lumpable" is the math word for a particular relation among state codings, used in that Tarasov paper.  I don't think we want to get into defining it precisely.  Instead, we want Fig 1 to make it intuitively clear, to Spencer and everyone else.  Probably need to add to its caption.
